---
title: "<center>**Selected dplyr functions in R**</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base", 'knitr', 'rmarkdown', 'tidyverse', 'moderndive',
'broom', 'nycflights13')
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])
}
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(
  .packages(), packages
), 'bib/packages.bib')
```

# Introduction

We start this by converting Titanic table data to tibble. This is for easy manipulation with tidyverse packages. We also use `dim` function to check the number of rows and columns of the tibble. 

```{r}
titanic_tibble <- tibble(data.frame(Titanic))
dim(titanic_tibble)
```

Checking the total number of passangers that boarded the Titanic.

```{r}
number_passengers <- titanic_tibble |> summarise(number_passengers = sum(Freq))
number_passengers
```

Checking the total number of passengers who succumbed and those who survived the ordeal.

```{r}
number_survived <- titanic_tibble  |>  
  group_by(Survived) |>
  summarise(Survival = sum(Freq))
number_survived
```
Selecting a few columns from the tibble.

```{r}
data_sex_age_freq <- titanic_tibble |> 
  select(Sex, Age, Freq)
data_sex_age_freq
```

# Mutate

Mutate creates a column using some conditional statements on other existing columns. For example, creating a new column which is 10 times the Freq column.

```{r}
titanic_tibble_mutate <- titanic_tibble |> 
  mutate(Freq_10 = Freq * 10)
head(titanic_tibble_mutate)
```

Mutate normally adds the newly created column as the last in the tibble. However, that can be changed by specifying where you want it by .after or .before arguments. Here I use .after.

```{r}
titanic_tibble_mutate <- titanic_tibble |> 
  mutate(Freq_10 = Freq * 10, .after = 1)
head(titanic_tibble_mutate)
```

Here I use .before, hence making it even the first column.

```{r}
titanic_tibble_mutate <- titanic_tibble |> 
  mutate(Freq_10 = Freq * 10, .before = 1)
head(titanic_tibble_mutate)
```

.after and .before do not only take column index like 1, 2, 3...but also take actual column names like is the case with Age column below.

```{r}
titanic_tibble_mutate <- titanic_tibble |> 
  mutate(Freq_10 = Freq * 10, .after = Age) 
head(titanic_tibble_mutate)
```

Can have it located before Sex column.

```{r}
titanic_tibble_mutate <- titanic_tibble_mutate |> 
  relocate(Age, .before = Sex)  
head(titanic_tibble_mutate)
```

Or even after Sex.

```{r}
titanic_tibble_mutate <- titanic_tibble_mutate |> 
  relocate(Age, .after = Sex)
head(titanic_tibble_mutate)
```

# Mutate and case_when

A interesting and important functionality which can be appended to `mutate` is called `case_when`. This is almost working like `if_else`.

```{r}
titanic_tibble |> mutate(Remarks = case_when(Survived == 'Yes' ~ 'Awesome, you made it', Survived == 'No' ~ 'RIP', TRUE ~ 'Wishing you well')) |> group_by(Remarks) |> summarise(Number = sum(Freq))
```

# Filter

Picks rows/records of a column meeting some defined conditions.

```{r}
titanic_female <- titanic_tibble |> 
  filter(Sex == 'Female') 
head(titanic_female) 
```

```{r}
titanic_tibble |> 
  dplyr::filter(Age == 'Child' & Sex == 'Female' & Survived == 'Yes' & Class == '1st')
```

```{r}
titanic_tibble |> 
  dplyr::filter(Age == 'Child' & Sex == 'Male' & Survived == 'Yes' & Class == '1st')
```

# Arrange

```{r}
titanic_df <- base::data.frame(Titanic) 
nrow(titanic_df)
```

```{r}
titanic_tibble <- tibble::tibble(titanic_df)
```

```{r}
titanic_tibble <- titanic_tibble |> 
  dplyr::arrange(Freq)
```


```{r}
utils::head(titanic_tibble) 
```


```{r}
utils::tail(titanic_tibble)
```

```{r}
titanic_tibble <- titanic_tibble |> 
  arrange(desc(Freq)) 
```


```{r}
head(titanic_tibble)
```


```{r}
titanic_tibble |> 
  arrange(desc(Freq), Survived)
```

# Correlation

```{r}
utils::data(flights)
```


```{r}
flights |> 
  dplyr::summarise(delay_cor = stats::cor(dep_delay, arr_delay))
```


```{r}
flights |> 
  dplyr::summarise(delay_cor = stats::cor(dep_delay, arr_delay, use = 'complete.obs'))
```


```{r}
flights |> 
  dplyr::select(dep_delay, arr_delay) |>
  stats::na.omit() |> 
  dplyr::summarise(delay_cor = stats::cor(dep_delay, arr_delay))
```

# Iris

```{r}
utils::data("iris")
```


```{r}
score_model <- stats::lm(Sepal.Length ~ Petal.Width, data = iris) 
score_model|> 
  base::summary()
```

```{r}
broom::tidy(score_model)
```


```{r}
broom::tidy(score_model) |>
  ggplot2::ggplot() + 
  ggplot2::geom_col(ggplot2::aes(x = term, y = statistic, fill = term))
```


```{r}
iris |> 
  dplyr::select(Sepal.Length, Petal.Width) |> 
  stats::lm() |> 
  broom::tidy() |> 
  ggplot2::ggplot() + 
  ggplot2::geom_col(ggplot2::aes(x = term, y = std.error, fill = term))
```


```{r}
broom::glance(score_model)
```


```{r}
moderndive::get_regression_summaries(score_model) 
```


```{r}
broom::augment(score_model)
```


```{r}
moderndive::get_regression_points(score_model)
```


```{r}
anova_iris <- stats::aov(Sepal.Length ~ Petal.Width, data = iris)
broom::tidy(anova_iris)
```


```{r}
utils::data(iris) 
utils::head(iris)
```


```{r}
stats::aggregate(x = base::c(iris$Sepal.Length), 
          by = base::list(iris$Species), 
          FUN = sum)
```


```{r}
iris |> 
  dplyr::group_by(Species) |> 
  dplyr::summarise(x = base::sum(Sepal.Length))
```

```{r}
iris |> 
  dplyr::group_by(Species) |> 
  dplyr::summarise(`sum(Sepal.Length)` = base::sprintf('%0.1f', base::sum(Sepal.Length)))
```

```{r}
iris |> 
  dplyr::group_by(Species) |> 
  dplyr::summarise(`sum(Sepal.Length)` = base::sprintf('%0.4f', base::sum(Sepal.Length)))
```

```{r}
library(reticulate)
```


```{python}
import numpy as np
my_python_array = np.array([2,4,6,8])
for item in my_python_array:
  print(item)
```

# References
