---
title: "<center>**k-means on rasters in R**</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base", 'knitr', 'rmarkdown',
'dismo',
              'raster',
              'ecbtools')
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])
}
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(
  .packages(), packages
), 'bib/packages.bib')
```

# Introduction

Here I demonstrate how to generate k-means output from stacked raster data in `R`. For the .Rmd behind this output, right click, copy the link and paste in your browser my [gitHub](https://github.com/Wycology/repoforrstudio/blob/main/kmeans.Rmd) account.

# Getting the data sets

I will make use of worldclim data and global administrative units data of Kenya.

```{r}
my_rasters <- getData('worldclim', res = 10, var = 'tmin')
my_polygon <- getData('GADM', country = 'KEN', level = 0)
masked_rasters <- mask(crop(my_rasters, my_polygon), my_polygon)
plot(masked_rasters[[2]])
```

# Stacking the rasters

```{r}
stacked_rasters <- raster::stack(masked_rasters) 
```

# Running k-means

Running k-means on raster data requires use of `raster.kmeans` function in `ecbtools` package. To install the `ecbtools` package use:


>remotes::install_github("ozjimbob/ecbtools")

You may need to install `remotes` package first to have the `install_github` function.

```{r}
kmeans_layer <- ecbtools::raster.kmeans(x = stacked_rasters, # The stack
                             k = 4,# Number of clusters
                             nstart = 3, # random sets 
                             geo = TRUE, # weighting by location (x,y coords)
                             geo.weight = 1) # wweight to location

raster::plot(kmeans_layer)
```

# Saving the raster output

Saving the raster into the current working directory for future reference or subsequent analysis/sharing as need may be.

```{r saving_output}
raster::writeRaster(kmeans_layer, 'output_clusters.tif', overwrite = TRUE) 
```

# References
