---
title: "**Mathematical equations in R Markdown**"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base",'knitr','rmarkdown')
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])
}
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(
  .packages(), packages
), 'bib/packages.bib')
```


# Mathematics Equations

Writing mathematical equations can be a daunting endeavor in R Markdown. Here I provide them as provided [here](https://rpruim.github.io/s341/S19/from-class/MathinRmd.html)

$x = y $


$x < y $

$x > y $

$x \le y $

$x \ge y $

$x^{n}$

$x_{n}$

$\overline{x}$

$\hat{x}$

$\tilde{x}$

$\frac{a}{b}$

$\frac{\partial f}{\partial x}$

$\displaystyle \frac{\partial f}{\partial x}$

$\binom{n}{k}$

$x_{1} + x_{2} + \cdots + x_{n}$

$x_{1}, x_{2}, \dots, x_{n}$

\mathbf{x} = \langle x_{1}, x_{2}, \dots, x_{n}\rangle$ 

$x \in A$

$|A|$

$x \in A$

$x \subset B$

$x \subseteq B$

$A \cup B$

$A \cap B$

$X \sim {\sf Binom}(n, \pi)$ 

$\mathrm{P}(X \le x) = {\tt pbinom}(x, n, \pi)$ 

$P(A \mid B)$

$\mathrm{P}(A \mid B)$ 

$\{1, 2, 3\}$

$\sin(x)$

$\log(x)$

$\int_{a}^{b}$

$\left(\int_{a}^{b} f(x) \; dx\right)$

$\left[\int_{-infty}^{\infty} f(x) \; dx\right]$

$\left. F(x) \right|_{a}^{b}$

$\sum_{x = a}^{b} f(x)$

$\prod_{x = a}^{b} f(x)$

$\lim_{x \to \infty} f(x)$

$\displaystyle \lim_{x \to \infty} f(x)$

# Greek Letters

$\alpha A$	

$\nu N$

$\beta B$	

$\xi\Xi$

$\gamma \Gamma$	

$o O$ (omicron)

$\delta \Delta$	

$\pi \Pi$

$\epsilon \varepsilon E$	

$\rho\varrho P$

$\zeta Z \sigma \,\!$	

$\sigma \Sigma$

$\eta H$	

$\tau T$

$\theta \vartheta \Theta$	

$\upsilon \Upsilon$

$\iota I$

$\phi \varphi \Phi$

$\kappa K$	

$\chi X$

$\lambda \Lambda$	

$\psi \Psi$

$\mu M$	

$\omega \Omega$

$\sum_{n=1}^{10} n^2$

$$\sum_{n=1}^{10} n^2$$

# Align equation

\begin{align*}
a & = b \\
X &\sim {\sf Norm}(10, 3) \\
5 & \le 10
\end{align*}

# References
