---
title: "<center>**Working with orange data in R**</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com) </center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
pacman::p_load(base, knitr, rmarkdown, datasets,
              'tidyverse',
              'parsnip',
              'gapminder',
              'ggstatsplot')
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(
  .packages(), packages
), 'bib/packages.bib')
```

# Checking data

```{r orange}
attach(Orange)

str(Orange)

orange_data <- Orange

orange_data <- as.data.frame(orange_data)

str(orange_data)

head(orange_data) 
```

# Some data checks

```{r}
orange_data %>% dplyr::select(age) %>% nrow()
orange_data_lm <- lm(circumference ~ age, data = orange_data) 
summary(orange_data_lm) %>% tidy()
```

# Simple plot

```{r}
plot(circumference ~ age, # Because we already attached the data in line 45
     col = "purple",
     pch = 19,
     cex = 2,
     xlab = "Age of orange (days)",
     ylab = "Circumference of orange (mm)",
     main = "Relationship between age and cricumference of oranges trees")
abline(orange_data_lm, lwd = 4, lty = "dotted", col = "magenta") 
text(300, 200, "R^2 = 0.8345\nP < 0.0001\ny = 17.4 + 0.107x")
```

# Checking largest circumference

```{r}
orange_data %>% dplyr::select(circumference) %>% max()
```

# Age of oldest tree in years

```{r}
orange_data %>% dplyr::select(age) %>% max()/365 # Assuming a year has 365 days
```

# Back to diamonds dataset

```{r}
diamonds %>% mutate(color2 = case_when(price > 400 ~ "High", price < 400 ~ "Moderate", TRUE ~ "Low")) %>% head()
```

# Checking speed of processing 

```{r}
start_time <- Sys.time()
head(gapminder)
gapminder %>% 
        dplyr::select(country, year, pop) %>% 
        filter(country == "Kenya") %>% 
        head(4)

end_time <- Sys.time()

duration2 <- round(end_time - start_time, 2)
duration2
```

# Histogram

```{r}
ggplot(data = gapminder, aes(pop)) + geom_histogram(col = "cyan", fill = "purple")
```


# Further histogram

```{r}
ggplot(gapminder, aes(log(pop))) + geom_histogram(col = "cyan", fill = "purple")

```


# Boxplot

```{r}
ggplot(gapminder, aes(continent, lifeExp, fill  = continent)) + geom_violin()
```

# Embracing ggstatsplot

```{r}
ggbetweenstats(data = movies_long, x = mpaa, y = rating)
```

# More steps further

```{r}
ggbetweenstats(data = movies_long, x = mpaa, y = rating, type = 'np', pairwise.display = 'all', p.adjust.method = 'fdr')
```

# Tagging outliers

```{r}
ggbetweenstats(data = movies_long, x = mpaa,y = rating, type = 'np', pairwise.display = "all",
        p.adjust.method = "fdr", outlier.tagging = TRUE, outlier.label = title, outlier.coef = 2)
```

## gghistostats

```{r}
gghistostats(data = movies_long, x = budget, test.value = 30)
```

## ggdotplotstats

```{r}
ggdotplotstats(data = movies_long, x = budget, y = genre, test.value = 52, test.value.line = TRUE)
```

## ggscatterstats

```{r}
ggscatterstats(
        data = movies_long,
        x = budget,
        y = rating,
        type = 'r',
        label.var = title,
        label.expression = budget > 150
        & rating > 7.5,
        marginal.type = 'density' 
)
```

## ggcormat

```{r}
ggcorrmat(dplyr::starwars)
```



## ggcormat as dataframe

```{r}
ggcorrmat(
        data = dplyr::select(msleep, sleep_rem, awake, brainwt),
        type = 'bf',
        output = 'dataframe'
)
```

## ggpiestats

```{r}
ggpiestats(
        data = dplyr::filter(.data = movies_long,
                             genre %in% c("Drama", "Comedy")),
        x = mpaa,
        y = genre
)
```

```{r}
ggpiestats(
        data = as.data.frame(Titanic),
        x = Class,
        counts = Freq,
        label = "both"
)
```

## ggbarstats

```{r}
ggbarstats(
        data = movies_long,
        x = genre,
        y = mpaa,
        package = 'ggsci',
        palette = 'default_igv'
)
```


# Extract coefficients

```{r}
mod <- stats::aov(
        formula = rating ~ mpaa * genre,
        data = movies_long
)
ggcoefstats(mod)
```

## grouped ggpiestats

```{r}
grouped_ggpiestats(
        data = mtcars,
        x = cyl,
        grouping.var = am
)
```

This is also available for grouped_ggdotplotstats, grouped_ggbarstats, grouped_ggscatterstats, grouped_gghistostats, grouped_ggbetweenstats, grouped_ggwithinstats, and grouped_cormat

## CUSTOMIZING THE PLOTS TO MY LIKING

```{r}
ggbetweenstats(
        data = movies_long,
        x = mpaa, 
        y = rating,
        results.subtitle = FALSE,
        ggtheme = hrbrthemes::theme_ipsum_tw(),
        palette = 'Darjeeling2',
        package = 'wesanderson'
)
```


## More playing

```{r}
ggbetweenstats(
        data = mtcars,
        x = am,
        y = wt,
        type = 'bf'
) +
        scale_y_continuous(sec.axis = dup_axis())
```

## Extracting only the useful statistics from ggplot using ggstatsplot

```{r}
results <- 
        ggstatsplot::ggpiestats(
                data = Titanic_full,
                x = Survived,
                y = Sex,
                output = 'subtitile'
        )
ggiraphExtra::ggSpine(
        data = Titanic_full,
        aes(x = Sex, fill = Survived),
        addlabel = TRUE,
        interactive = FALSE
) +
        labs(subtitle = results)
results
```
